---
title: "Regression"
author: "Your name"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Prerequisite

```{r}

rm(list = ls()) # clean environment

# install.packages("stargazer")

library(stargazer)
library(tidyverse)

# load data
df <- read_csv(file = "data/sim.csv")
```


# Regression analysis

Load the dataset **sim.csv**, examine the correlation matrix of the numerical variables, and create a scatterplot with the y and x variables using **ggplot**. Next, apply linear regression using `geom_smooth()` with the argument `se=FALSE`.

```{r}

# correlation


# scatter plot


# scatter + linear fit



```


Run a regression of y on x and use the `stargazer()` function to report the estimated table. 

**Note**: remember to set the argument type to choose whether to print the results in \LaTeX (for PDF) or *"text"* (for R console). Also, if you want to knit the document in PDF, make sure to activate the code chunk option `results='asis'`.

```{r,results='asis'}


```


# Confounding

First, generate a **scatter plot** between x and y, visually distinguishing the data points based on the categorical variable z using color aesthetics. Then, construct two boxplots comparing the distributions of y and x with respect to the binary indicator variable z. 

*Note: z is a binary categorical indicator, so it's preferable to create a factor variable version of z to ensure proper mapping of dichotomous colors in ggplot.*

```{r}


# scatter plot of y on x, conditional on z


# boxplot of y on z, use the fill aesthetics


# boxplot of x on z, use the fill aesthetics



```

Now, conduct a regression of y on both x and z. Report the regression model for y with both x and z, and compare it to the previous model using stargazer.

```{r,results='asis'}



```


Visualize the regression line between y and x, this time conditioning on z. Simply run the same code as before, but include z in the color `aes`thetics.


```{r}

# scatter + linear fit, of y on x



# scatter + linear fit, of y on x conditional on z




```



# Extra: Prediction


*Note: This final section is optional and does not need to be submitted for earning participation points in this activity.*

To compare the slope of x between the two estimated models, we can calculate the predicted values of y for a range of hypothetical x values. To achieve this, we will create a `data.frame()` containing hypothetical x values ranging from 1 to 12 and then use the `predict()` function.

For the second model, which incorporates z as a covariate, we will need to generate another dataset of hypothetical values that includes the mean value of z. This ensures that the prediction is based on changes in x while maintaining z constant at its mean.


```{r}
# Predictions for first model

hyp_x <- data.frame(x=1:12) # hypothetical values

pred_m1 <- predict(m1, # estimated model 1
                   newdata=hyp_x)


# Predictions for second model

hyp_x2 <- data.frame(x=1:12,
                     z=mean(df$z))

pred_m2 <- predict(m2, newdata=hyp_x2)


# combine hypothetical values and predictions

pred_df <- cbind(hyp_x,pred_m1,pred_m2)

pred_df

```



# Extra code: visualize the predictions


```{r}

# first, pivot the data frame into long format

pred_df_long <-
  pivot_longer(data=pred_df,
             cols = c(pred_m1,pred_m2),
             names_to = "model",
             values_to = "fit")

ggplot(data=pred_df_long,
       mapping=aes(y=fit,
                   x=x,
                   color=model)) +
  geom_line() +
  theme_minimal()


```






```{r}

# notice that we can also create the same predictions by
# using the coef() function and use the parameters directly

coef(m1)

coef(m2)

# extract parameters from model 1:

intercept_m1 <- coef(m1)[1]
B1_m1 <- coef(m1)[2]


# extract parameters from model 2:

intercept_m2 <- coef(m2)[1]
B1_m2 <- coef(m2)[2]

# create data set with predictions using the extracted parameters:

data.frame(
  fit1 = intercept_m1 + B1_m1*1:12,
  fit2 = intercept_m2 + B1_m2*1:12
)


```
